---
title: "Machine Learning 1"
author: "Irmak Ozarslan"
date: "2025-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis of Economic Indicator's Relationship with Leader Idealogies

#### **Table of Contents**


## **1. Introduction**



Our research questions are the following :

**Are economic indicators good predictors of leader ideologies?**

**Can we predict military spending from leader ideologies and other economic indicators?**

\vspace{5cm}

## **2. Data**
\vspace{3cm}

### **2.1. Data Source**





### **2.2. Description of the Data Source:**
\vspace{5cm}

#### **Data 1: Global Economic Indicators :**


#### **Data 2: Leader Ideologies :**

#### **Data 3: Military Spending by Country :**


### **3.2 Data Merging**

-   Merging all the 4 tables was done with the Columns Year and Country.Name

-   We did visual inspection of Country.Name in different datasets, if not matched it was standardized. However, since all datasets are based on ourworldindata, the Country.Name's were identical and there were missing columns only due to missing data and not mismatch of names.





```{r}
#LIBS:
library(countrycode)
library(dplyr)
library(ggplot2)
library(magrittr)
library(readr)
library(rmarkdown)
library(tidyr)
library(tidyverse)
library(tinytex)
library(knitr)
library(shiny)
library(sf)
library(dplyr)
library(scales)
library("GGally")
library(contrast)
library(effects)
library(MCMCglmm)
library(languageR) 
library(lmerTest)
library(coefplot)
library(stats)
library(curl)
library(data.table)
library(car)
library(phia)
library(MuMIn)
library(glmmTMB)

```



### **4. Visual Inspection of given Data**
\vspace{5cm}

```{r}
#PLOT EDA CORRS
```


## 5. **Hypothesis:**

This study investigates the relationship between mental health issues, unemployment rates, with suicide rates and drug-related deaths across countries over the past 19 years.These variables are measured as the percentage of the population, across most of the countries in the world, from the years 2000 to 2019.

 The final dataset includes the following variables:



\vspace{5cm}

#### **Hypothesis 1:** We would expect that higher the mental health disorder percentages and unemployment, there will be higher suicide rates.

#### **Hypothesis 2:** We would expect to find a positive relationship between mental health conditions, drug overdoses and unemployment.

\vspace{5cm}

### **5.1 Modeling:*


### LINEAR MODELS:

$$

Where:

Suicide Rates are the dependent variable (the rate of suicide in country$_i$

Schizophrenia Rates, Depressive Disorder Rates, Anxiety Disorder Rates, Unemployment Rates are the independent variables.

-   $\beta_0$ is the intercept term

-   $\beta_1$, $\beta_2$... $\beta_7$ are the coefficients representing how each predictor (e.g., schizophrenia rates, unemployment) affects suicide rates

-   $\epsilon_i$ is the error term.

    To address the third hypothesis, we model the relationship between drug overdoses, unemployment rates, mental health disorder rates. The GLM equation for drug overdoses in relation to multiple predictors is specified as:

$$




### **6.2 Generalized Linear Modelling (Multivariate Analysis)**

\vspace{5cm}

#### **6.2.1: Logistic Regression:**

#### **6.2.1: poisson Regression:**

#### **6.2.1: binary Regression:**

#### **6.2.1: GAM:**

#### **6.2.1: SVM:**

#### **6.2.1: NN:**





